---
title: "Play with RMD"
output: html_document
date: 3 Dec 2024
author: Javier Magide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r scatter}
plot(1:100)
```

```{r download}
download.file(url="https://raw.githubusercontent.com/cambiotraining/reproducibility-training/master/data/gapminder_data.csv", destfile="data/gapminder_data.csv")
```

```{r plot_hist, echo=FALSE}
hist( rnorm(1000), col='red')
```

```{r load_data, message=FALSE}
#load tidyverse library
library(tidyverse) # used for data manipulation
library(rmarkdown) # used for paged_table function
library(kableExtra) # used for table
library(ggpubr) #used for ggarrange function 

#read file into R
pop_data <- read_csv("data/gapminder_data.csv")

#create a table with data from European countries in 2007 showing the countries 
#with the largest life expectancy at the top 
euro_data_tbl <- pop_data %>% 
                    filter(continent == "Europe" & year == 2007) %>% 
                    select(-continent, -year) %>% 
                    arrange(desc(lifeExp)) %>% 
                    rename(Country = country, "Population Size" = pop, 
                      "Life Expectancy" = lifeExp, "GDP" = gdpPercap)


```

## Table

```{r kbl}
euro_data_tbl %>% 
          kbl() 
```

```{r kbl_beautify}
euro_data_tbl %>% 
          kbl() %>%
          kable_styling(bootstrap_options = c("striped", "hover")) 
```

```{r paged-table}
paged_table(euro_data_tbl)
```

## Challenge

Next, the life expectancy in Europe is observed across different years.  The aim of this is to check if there was a change in life expectancy over the years.  The data used for the life expectancy over the years analysis is the gapminder data and the following data manipulation was performed:

```{r fig-data}
#keep on European data and change year to factor
euro_data_fig <- pop_data %>% 
                    filter(continent == "Europe")  %>% 
                    mutate(year=as_factor(year))

#keep only United Kingdom data and change year to factor
uk_data_fig <- pop_data %>%
                    filter(country == "United Kingdom") %>%
                    mutate(year=as_factor(year))

```


The life expectancy for Europe over the years is plotted as following:

```{r fig, warning=FALSE, fig.cap="Life Expectancy in Europe over the years"}
#Euro plot
euro_plot <- euro_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_violin() +
                stat_summary(fun.y = median, geom = "point")

#draw euro plot
euro_plot

```

\ 

Comparison of life expectancy from Europe and United Kingdom:

```{r fig-comp, warning=FALSE, fig.cap="Life Expectancy in European and UK over the years"}
#UK plot
uk_plot <- uk_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_point()

#draw euro plot next to UK plot
ggarrange(euro_plot, uk_plot, ncol=1, nrow=2, labels="AUTO")
```

